---
title: "course_project"
author: "Luyi Tian"
date: "26 July 2015"
output: html_document
---
##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 



##Data 


The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

##What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 
##prepare the library
```{r}
library(caret)
library(randomForest)
```
##load the data
```{r}
#load data and read the csv file
trainURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
traindata <- read.csv(url(trainURL), na.strings=c("NA","#DIV/0!",""))
testdata <- read.csv(url(testURL), na.strings=c("NA","#DIV/0!",""))

#look at the data
dim(traindata)
dim(testdata)

# Delete columns with missing values
traindata <- traindata[,colSums(is.na(traindata)) == 0]
testdata <- testdata[,colSums(is.na(testdata)) == 0]

# Deleting irrelevant variables like user_name, raw_timestamp.
traindata <- traindata[,-c(1:7)]
testdata <- testdata[,-c(1:7)]
```
#partition training data - cross validation
divide the data into two part, first part as training data and second as testing data.
```{r}
ran_ind <- createDataPartition(y=traindata$classe, p=0.66, list=FALSE)
Training <- traindata[ran_ind, ]
Testing <- traindata[-ran_ind, ]
dim(Training)
dim(Testing)
```
#machine learning - random forest
I use the random forest algorithm since random forest is considered as one of the top ML algorithms. 
```{r}
Fit <- randomForest(classe ~. , data=Training)
# Predicting on test data we created
predictions <- predict(Fit, Testing, type = "class")
confusionMatrix(predictions, Testing$classe)
```
The expected out-of-sample error is calculated as 1 - accuracy for predictions made against the cross-validation set. Since our testing data set comprises of only 20 cases, we can expect that very few of the test samples will be missclassified.
```{r}
# Testing results on Testing data set
predictions2 <- predict(Fit, testdata, type = "class")
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions2)
```
